---
title: "HW 3"
author: "Genavieve Middaugh"
date: "2025-02-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The theories
People have a lot of pet theories about what explains the variation in prices between gas stations. Here are
several such theories:
A) Gas stations charge more if they lack direct competition in sight.
B) The richer the area, the higher the gas prices.
C) Gas stations at stoplights charge more.
D) Gas stations with direct highway access charge more.
E) Shell charges more than all other non-Shell brands.

Which of these theories seem true, and which are unsupported by data? Take each theory one
by one and assess the evidence for the theory in this data set.
Your discussion of each theory should include three mini-sections:
- Claim: a statement of the theory itself.
- Evidence: the evidence for or against the theory, in the form of any relevant numerical and/or visual
summaries If the theory looks correct, provide an estimate of the effect size: that is, how large is the difference
(e.g. for highway vs. non-highway gas stations) and/or association (e.g. between income and price)? If the
theory is unsupported by the data, explain why.
- Conclusion: your conclusion about whether the theory is supported or unsupported by the data.
No single theory should require more than a single typed page to assess, including any figures, tables, or
numbers you use to make your case. Less than 1 page per theory is perfectly acceptable—indeed, preferable,
as long as you can cover all the bases.
In assessing the evidence for each theory, make sure you appropriately deal with a major issue: uncertainty.
Remember, this is just a sample of gas stations, and there is always uncertainty when generalizing from a
specific sample to a wider population. Therefore, you need to quote confidence intervals rather than just
single-number estimates for these differences. So, for example, it’s not enough to say something like, “the
difference in price between gas stations on and off the highway is X cents,” and to draw your conclusion from
this statement. Instead, you have to say something like, “the difference in price between gas stations on an

```{r}
gasprices <- read.csv("gasprices.csv")

```

off the hig